---
title: "Plot Replication"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  pdf_document:
    highlight: "kate"
    df_print: "kable"
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

\newcommand{\rot}[1]{
\begin{bmatrix}
\cos{(#1)} & -\sin{(#1)}\\

\sin{(#1)} & \cos{(#1)}\\
\end{bmatrix}
}
```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  warning = FALSE,
  tidy = TRUE,
  fig.align = "center",
  highlight = TRUE
)

library(tidyverse)
library(rlang)
```

This is an attempt to replicate the plot found (here)["https://www.economist.com/britain/2021/07/10/some-britons-crave-permanent-pandemic-lockdown"].
```{r}
data <- tibble(
  question = factor(
    rep(1:7, each = 3),
    labels =
      c(
        "Masks in shops and on public transport",
        "Ten-day quarantine on returning from abroad",
        "Foreign travel only with proof of vaccination",
        "Social distancing in theaters, pubs, and sports grounds",
        "Check-in for contact tracing purposes in pubs and restaurants",
        "Nightclubs and casinos closed",
        "A 10pm curfew"
      )
  )
  ,
  opinion = factor(
    rep(1:3, times = 7),
    labels = c(
      "For a month after July 19th",
      "Until covid-19 is under control globally",
      "Permanently, regardless of covid-19"
    )
  ),
  value = c(replicate(7, c(60, 50, 40) + rnorm(
    3, mean = 5, sd = 2.5
  )))
)

```


I couldn't figure out how to replicate the shaded rectangles under the axis labels, though I feel it turned out well otherwise.
```{r}
library(ggplot2)
library(forcats)

breaks <- seq(0, 70, by = 10)
ggplot(data, aes(
  x = fct_rev(question),
  y = value,
  fill = fct_rev(opinion)
)) +
  geom_col(position = "dodge", width = .5) +
  discrete_scale(aesthetics = "fill", scale_name = "economist", palette = function(x) rev(c("#8FC2D1", "#3DBFC4", "#045589")), expand = expansion(mult = c(0, 0))) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0)),
    breaks = breaks,
    labels = breaks,
    limits = c(0, 75),
    position = "right"
  ) +
  coord_flip() +
  labs(
    caption = c("Source: Ipsos Mori", "*Making listed restrictions compulsory"),
    title = "Sealing clubs",
    subtitle = "Britain, covid restrictions, % who support*",
    fill = NULL,
    x = NULL,
    y = NULL
  ) +
  guides(fill = guide_legend(override.aes = list(stroke = "#ADD3DE"), reverse = TRUE)) +
  theme(
    axis.line.y = element_line(color = "black", size = .5),
    axis.ticks = element_blank(),
    axis.text.x = element_text(face = "bold", color = "#2C313B"),
    axis.text.y = element_text(hjust = 0, color = "#2C313B"),
    legend.background = element_rect(fill = "#ADD3DE"),
    legend.direction = "vertical",
    legend.position = c(1, 1.1),
    legend.justification = c(1, 0.2),
    legend.margin = margin(),
    legend.key = element_rect(fill = "#ADD3DE"),
    legend.key.width = unit(.65, "lines"),
    legend.key.height = unit(.65, "lines"),
    panel.background = element_rect("#ADD3DE"),
    plot.background = element_rect(fill = "#ADD3DE"),
    panel.grid = element_line(color = "grey", size = .5),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = unit(c(1.5, 1, 0.5, 0.5), "lines"),
    plot.caption = element_text(hjust = c(-1.55,1), color = "#2C313B"),
    plot.subtitle = element_text(vjust = 1),
    plot.title = element_text(vjust = 1, face = "bold"),
    plot.title.position = "plot"
  )
```



