---
title: "`r params$title`"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
header-includes:
  - \setlength{\parindent}{2em}
  - \setlength{\parskip}{2em}
params:
    title: "New York Times Bestsellers Analysis"
output:
    github_document:
      df_print: "kable"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  fig.path = file.path(
    "../figure",
    tools::file_path_sans_ext(
      knitr::current_input()
    ),
    "/"
  ),
  message = FALSE,
  tidy = "styler",
  warning = FALSE,
  fig.align = "center",
  highlight = TRUE
)
```

```{r}
library(tidyverse)

nyt_titles <- read_tsv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-10/nyt_titles.tsv")
nyt_full <- read_tsv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-10/nyt_full.tsv")
```

# Largest Rank Gaps 

Which books had the longest gaps between 
debt rank and best rank?
```{r}
nyt_full |>
  group_by(title_id, title) |>
  summarize(
    rank_change = max(rank) - rank[which.min(week)],
    duration = cut(n(),
      breaks = c(0:9, seq(10, 100, by = 10), Inf),
      labels = as.character(
        c(
          1:10, paste(seq(10, 90, by = 10), seq(20, 100, by = 10), sep = "-"),
          "100+"
        )
      ),
      right = TRUE
    )
  ) |>
  ggplot(aes(x = rank_change)) +
  geom_bar(aes(fill = duration))
```

# Longest Reigns

Which books had the longest reigns?
Most are familiar, but a few have been forgotten. _The Robe_'s reign was not contiguous. 
```{r}
nyt_full |>
  group_by(title) |>
  summarize(length = n(), entrance = min(week), exit = max(week)) |>
  slice_max(n = 10, order_by = length)
```

So which had the longest contiguous reigns?
```{r}
nyt_full |>
  group_by(title) |>
  summarize(longest_streak = with(rle(
    seq.Date(
      min(week),
      max(week),
      by = "weeks"
    ) %in%
      week
  ), max(lengths[values]))) |>
  slice_max(n = 10, order_by = longest_streak)
```

# Most Represented Authors

Checking the most represented authors turns 
up some of the usual suspects. 
```{r}
top_authors <- nyt_full |>
  distinct(title_id, .keep_all = TRUE) |>
  mutate(author = str_remove(author, "^.\\sby\\s")) |>
  count(author) |>
  slice_max(n = 10, order_by = n)

top_authors
```


```{r}
nyt_full |>
  filter(author %in% top_authors[["author"]]) |>
  ggplot(aes(x = week, y = rank, color = author)) +
  geom_line(aes(group = as.factor(title_id))) +
  facet_wrap(. ~ author, ncol = 2) +
  theme(legend.position = "bottom")
```

# Title Worlds

The most common title words paint something of a grim picture.
```{r}
library(tidytext)

top_words <- nyt_full |>
  distinct(title_id, .keep_all = TRUE) |>
  unnest_tokens(word, title) |>
  anti_join(stop_words, by = "word")

top_words |>
  count(word) |>
  arrange(desc(n))
```


```{r}
top_words |>
  semi_join(top_authors,
    by = "author"
  )
```


```{r}
author_words <- top_words |>
  group_by(author) |>
  summarize(word = paste(word, collapse = " "))
```


```{r}
library(quanteda)
DFM <- tokens(author_words[["word"]]) |>
  setNames(author_words[["author"]]) |>
  dfm() |>
  dfm_tfidf(scheme_tf = "prop", base = exp(1))
```

Tf-idfs aren't as distinctive as I hoped for every author, but some are instantly recognizable.
```{r}
tidy(DFM) |>
  rename(author = document) |>
  mutate(author = str_remove(author, "^.\\sby\\s")) |>
  filter(author %in% top_authors[["author"]]) |>
  group_by(author) |>
  slice_max(n = 10, order_by = count, with_ties = FALSE) |>
  mutate(term = fct_rev(reorder_within(term, count, author, FUN = mean))) |>
  ggplot(aes(x = term, y = count, fill = term)) +
  geom_col() +
  scale_x_reordered() +
  facet_wrap(. ~ author, ncol = 2, scales = "free") +
  labs(y = "Log tf-idf") +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, size = 9)
  )
```
