---
title: "Notes"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  pdf_document:
    highlight: "kate"
    df_print: "kable"
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

\newcommand{\rot}[1]{
\begin{bmatrix}
\cos{(#1)} & -\sin{(#1)}\\

\sin{(#1)} & \cos{(#1)}\\
\end{bmatrix}
}
```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  warning = FALSE,
  tidy = TRUE,
  fig.align = "center",
  highlight = TRUE
)

library(tidyverse)
library(rlang)
```


```{r}
library(tidytuesdayR)
library(ggplot2)

dubois <-  tt_load(2021, week = 8)
```

```{r}
freed_slaves <- dubois$freed_slaves
head(freed_slaves)
```

```{r}
labs <- sprintf("%.0f%%", rep(dubois$freed_slaves$Free, each = 2))
labs_coords <-
  c(rep(dubois$freed_slaves$Slave[-nrow(dubois$freed_slaves)], each = 2),
    rep (dubois$freed_slaves$Slave[nrow(dubois$freed_slaves) - 1], 2))
p1 <-
  freed_slaves %>% pivot_longer(c(Slave, Free), values_to = "Number", names_to = "Status") %>%
  ggplot(aes(
    x = Year,
    y = Number,
    fill = Status,
    alpha = Status
  )) +
  stat_summary(geom = "area",
               position = "fill",
               fun = "identity") +
  geom_vline(
    aes(xintercept = Year),
    color = "black",
    size = .5,
    alpha = .5
  ) +
  scale_fill_manual(values = c(Free = "green4", Slave = "Black")) +
  scale_alpha_manual(values = c(Free = .9, Slave = 1)) +
  scale_x_continuous(
    expand = c(.025, 0),
    breaks = seq(1790, 1870, by = 10),
    position = "top"
  ) +
  scale_y_continuous(expand = c(0, 0), labels = scales::percent_format()) +
  annotate(
    "text",
    x = 1830,
    y = .95,
    label  = "FREE - LIBRE",
    fontface = "bold",
    size = 5
  ) +
  annotate(
    "text",
    x = 1830,
    y = .6,
    "label" = "SLAVES\nESCLAVES",
    color = "white",
    size = 7,
    fontface = "bold"
  ) +
  geom_text(aes(y = (labs_coords / 100) + .025), label = labs) +
  labs(y = NULL, x = NULL) +
  theme(
    legend.position = "none",
    axis.ticks = element_blank(),
    axis.text.x = element_text(size = 12, face = "bold"),
    axis.text.y = element_blank(),
    panel.background = element_rect(fill = "white")
  )
ggsave("dubois1.png", plot = p1, path = "../outputs/plots")
```
```{r}
factor_levels = c(
  "Agriculture, Fisheries and Mining",
  "Domestic and Personal Service",
  "Manufacturing and Mechanical Industries",
  "Trade and Transportation",
  "Professions"
)
gap <- 20
occupation <- dubois$occupation %>% group_by(Group) %>%
  mutate(
    Occupation = factor(Occupation,
                        levels = factor_levels,
                        ordered = TRUE),
    ord = order(Occupation, factor_levels)
  ) %>%
  arrange(ord) %>%
  mutate(
    xmax = gap + cumsum(Percentage * ((100 - (
      gap * 2
    )) / 100)) + (100 * (cur_group_id() - 1)),
    xmin = lag(xmax, default = gap  + (100 * (cur_group_id(
      
    ) - 1))),
    ymin = 0,
    ymax = 1,
    label_x = xmin + (xmax - xmin) / 2
  ) %>%
  ungroup()

key <-
  setNames(c("red", "yellow", "blue", "#DFBF9F", "gray"),
           levels(occupation$Occupation))
circle_data <- tibble(
  color = key,
  x = c(5, 90, 195, 110, 100),
  y = .9,
  size = 7,
)
text_data <- tibble(
  label = c(
    "AGRICULTURE, FISHERIES\n AND MINING",
    "DOMESTIC AND\n PERSONAL SERVICE",
    "MANUFACTURING AND\n MECHANICAL INDUSTRIES",
    "TRADE AND\n TRANSPORTATION",
    "PROFESSIONS"
  ),
  x = c(5, 90, 195, 110, 100),
  y = .8,
  size = 4.35,
  hjust = c("left", "right", "left", "right", "right")
)
group_labs <-
  tibble(
    x = c(50, 150),
    y = 1.1,
    label = c("NEGROES.", "WHITES.")
  )
p2 <- occupation %>%
  ggplot(
    aes(
      x = Percentage,
      xmin = xmin,
      xmax = xmax,
      ymin = ymin,
      ymax = ymax,
      group = Group,
      fill = Occupation
    )
  ) +
  geom_rect(position = "dodge", show.legend = FALSE) +
  geom_text(aes(
    label = paste0(Percentage, "%"),
    x = label_x,
    y = .9,
    
  ), size = 2.5) +
  coord_polar(clip = "off",
              start = -pi / 2) +
  xlim(c(0, 200)) +
  scale_fill_manual(values = key) +
  geom_point(
    data = circle_data,
    aes_auto(names(circle_data)),
    inherit.aes = FALSE,
    show.legend = FALSE
  ) +
  geom_text(
    data = text_data,
    aes_auto(names(text_data)),
    inherit.aes = FALSE,
    show.legend = FALSE
  ) +
  geom_text(data = group_labs, aes_auto(names(group_labs)), inherit.aes = FALSE) +
  scale_color_identity() +
  labs(title = "OCCUPATIONS OF NEGROES AND WHITES IN GEORGIA.") +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    panel.grid  = element_blank(),
    panel.background = element_rect(fill = "white"),
    plot.title.position = "plot",
    plot.title = element_text(hjust = .5, face = "bold")
  ) 

ggsave("dubois2.png", plot = p2, "../outputs/plots/")
```






