---
title: "Scraping Maryland Election Results"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  pdf_document:
    highlight: "kate"
    df_print: "kable"
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

\newcommand{\rot}[1]{
\begin{bmatrix}
\cos{(#1)} & -\sin{(#1)}\\

\sin{(#1)} & \cos{(#1)}\\
\end{bmatrix}
}
```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  warning = FALSE,
  tidy = TRUE,
  fig.align = "center",
  highlight = TRUE
)

library(tidyverse)
library(rlang)
```

For this exercise I'll scrape and tidy official 2020 election results for
Baltimore County, Maryland.
```{r}
library(rvest)
url <- "https://elections.maryland.gov/elections/2020/results/General/gen_results_2020_4_by_county_04-1.html"

selectors <- c("#primary_right_col h2", "h3", ".mdgov_OverflowTable .table-stripe")
funs <- list(html_text, html_text, html_table)


raw <- read_html(url)
scraped <- map(selectors, ~html_elements(x = raw, css = .x)) %>% 
  map2(., funs, ~rlang::exec(.fn = .y, .x))  

districts <- str_match(scraped[[2]], "District\\s(\\d)+")[,2]
inds <- which(complete.cases(districts))
districts <- districts[inds]

tidied <- scraped[[3]]
names(tidied)[inds] <-districts
```

Tidying is more involved then expected. I decide to drop a retention election,
since a yes/no vote can't easily be represented in a table among votes for 
candidates.
```{r}
races <- scraped[[1]]
races <-
  c(races[1:min(inds) - 1], rep("U.S. Congress", length(inds)), races[(min(inds) +
                                                                         2):length(races) - 1])
tidied <- tidied %>% discard( ~ names(.x)[1] != "Name") %>%
  map2(., races, ~ mutate(.x, across(c(
    where(is.character), -c(Party, Name)
  ), parse_number)) %>% 
    mutate(Race = .y)) %>%
  bind_rows(.id = "District")
```

My usual plotting theme.
```{r}
theme_standard <- ggplot2::theme(panel.background = element_blank(), 
        panel.border = element_rect(color = "black", fill = NA), 
        panel.grid = element_blank(), panel.grid.major.x = element_line(color = "gray93"), 
        legend.background = element_rect(fill = "gray93"), 
        plot.title = element_text(size = 15, family = "sans", 
            face = "bold", vjust = 1.3), plot.title.position = "plot", 
        plot.subtitle = element_text(size = 10, family = "sans"), 
        legend.title = element_text(size = 10, family = "sans", 
            face = "bold"), axis.title = element_text(size = 9, 
            family = "sans", face = "bold"), axis.text = element_text(size = 8, 
            family = "sans"), strip.background = element_rect(color = "black", 
            fill = "black"), strip.text.x = element_text(color = "white"), 
        strip.text.y = element_text(color = "white"))
    ggplot2::theme_set(theme_standard)
```

Making the plot. The diffeernces in relative proportions of voting mode by 
congressional district are interesting.
```{r}
tidied %>%  group_by(Race, District) %>% 
  summarize(across(c(`Early Voting`, `Election Day`, `By Mail`  ), ~ sum(.x)/sum(Total),
                   .names = "{.col} Proportion"), .groups = "drop") %>% 
  mutate(Race = if_else(Race == "U.S. Congress", sprintf("%s (District %s)", Race, District), Race)) %>% 
  select(-District) %>% 
  pivot_longer(-Race, names_to = "Voting Mode", values_to = "Proportion") %>% 
  ggplot(aes(x = `Voting Mode`, y = Proportion, fill = `Voting Mode`)) +
  geom_col(position = "stack") +
  scale_y_continuous(labels = scales::percent_format())+
  facet_wrap(Race ~., nrow = 2, ncol =3) +
  scale_fill_viridis_d()+
  labs(title = str_wrap("2020 Baltimore County Election Results by Mode of Voting",width = 40)) +
  theme(
        axis.text.x = element_blank(),
        legend.position = "bottom",
        plot.title = element_text(size = 10),
        strip.text = element_text(size = 6))
  
```

