---
title: "Tidying FAA Data"
author: "Ryan Heslin"
date: "`r  Sys.Date()`"
output: html_document
---

 <!-- Standard custom LaTeX commands -->
\newcommand{\abcd}{\begin{bmatrix}a&b\\
c&d\end{bmatrix}}
\newcommand{\m}[1]{\begin{bmatrix}#1\end{bmatrix}}

\newcommand{\vect}[1]{\begin{pmatrix}#1\end{pmatrix}}

\newcommand{\meq}[1]{\begin{split}#1\end{split}}

\newcommand{\bym}[1]{#1\times{m}}

\newcommand{\nby}[1]{n\times{#1}}

\newcommand{\subsp}[2]{\Bigg\{\begin{bmatrix}#1\end{bmatrix}:#2\Bigg\}}

\newcommand{\proj}[2]{\text{proj}_#1(#2)}

\newcommand{\refl}[2]{\text{refl}_#1(#2)}

\newcommand{\sumn}{\sum_{i=1}^n}

<!-- % 1: term 1 -->
<!-- % 2: subscript 1 -->
<!-- % 3: term 2 -->
<!-- % 4: subscript 2 -->
<!-- % 5. operation -->
\newcommand{\dotsn}[5]{#1_{1}#3_{1}#5{#1}_{2}#3_{2}{#5}\dots{#5}#1_{#2}#3_{#4}}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.pos = "", warning = FALSE,
                      tidy = TRUE, fig.align = "center")

library(tidyverse)
library(rlang)
```

This dataset is a nastily formatted .xslx table breaking down 2019 FAA pilot
licenses by region and type. To extract useful data, I'll have to remove useless
rows and extract information from cell formattting.
```{r}
library(openxlsx)
library(tidyxl)


formats <- xlsx_formats("../data/faa.xlsx")$local$font$bold
cells <-  xlsx_cells("../data/faa.xlsx") %>% 
  filter(sheet=="Table 3" & str_detect(address, "^A")  & local_format_id %in% which(formats) & between(row, 8, 59))
```

The raw dataset.
```{r}
faa_raw <- read.xlsx("https://www.faa.gov/data_research/aviation_data_statistics/civil_airmen_statistics/media/2019-civil-airmen-stats.xlsx", sheet =4,start = 6, na.strings = "N/Ap", cols = 1:11)
head(faa_raw)

str(faa_raw)
```

The hardest part is to identify the cells of the first column with bold 
formatting, indicating categories. I do this using the openxlsx package, then 
extract the category names from the relevant cells. That done, I use those
cells to fill the rest of the column. For simplicity's sake, I don't 
bother with subcateegoreis (e.g., commercial fixed-wing vs. helicopter licenses)

After that, I clean up column names and delete an uneccesary row.
```{r}
faa_cleaned <- faa_raw[1:53,] %>% 
mutate(category = NA_character_)
faa_cleaned[cells$row - 6, "category"] <- str_remove(cells$character, "\\s\\d.*|\\s?-.*") %>% 
  str_trim()

faa_cleaned <- faa_cleaned %>% fill(category, .direction = "down") %>% 
  rename_with(~str_to_lower(.x) %>% 
                str_replace_all("-?\\.", "_") %>% 
                str_remove_all("_[^a-zA-Z]+")
                ) %>% 
  filter(!str_detect(class_of_certificate, "Total|^Airplane")) %>% 
  mutate(str_remove(category, "\\sCetificates"))

```

From then on it's an easy pivoting job. 
```{r}
faa_cleaned <- faa_cleaned %>% 
  select(-total) %>% 
  pivot_longer(cols = alaskan:outside_u_s, names_to = "Region", values_to = "Licenses")
```

Now to actually plot the thing.
```{r}
faa_cleaned %>% 
  ggplot(aes(x = category, y = Licenses, fill = category)) +
  geom_col(position = "dodge") +
  facet_wrap(~Region) +
  theme(legend.position = "bottom", axis.text.x = element_blank(),
        legend.text = element_text(size = 8), legend.title = element_blank(),
        legend.background = element_rect(fill = "khaki4", color = "gray93"),
        plot.background = element_rect(fill = "cadetblue3", color = "gold"),
        strip.text = element_text(color = "white"), strip.background = element_rect(fill = "black"),
        panel.background = element_rect(fill = "white")) +
  labs(x = "Category", title = "2019 Pilots' Licenses by Region and Category") +
  guides(fill = guide_legend(override.aes = list(color = "black")))
```

