---
title: "`r params$title`"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
header-includes:
  - \setlength{\parindent}{2em}
  - \setlength{\parskip}{2em}
params:
    title: "FAA Data Analysis"
output:
    github_document:
      df_print: "kable"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  fig.path = file.path(
    "../figure",
    tools::file_path_sans_ext(
      knitr::current_input()
    ),
    "/"
  ),
  message = FALSE,
  tidy = "styler",
  warning = FALSE,
  fig.align = "center",
  highlight = TRUE
)
```

This dataset is a nastily formatted `.xslx` table breaking down 2019 FAA pilot
licenses by region and type. To extract useful data, I'll have to remove useless
rows and extract information from cell formatting.
```{r}
library(tidyverse)
library(openxlsx)
library(tidyxl)

formats <- xlsx_formats("../data/faa.xlsx")$local$font$bold
cells <- xlsx_cells("../data/faa.xlsx") %>%
  filter(sheet == "Table 3" & str_detect(address, "^A") & local_format_id %in% which(formats) & between(row, 8, 59))
```

The raw dataset.
```{r}
faa_raw <- read.xlsx("https://www.faa.gov/data_research/aviation_data_statistics/civil_airmen_statistics/media/2019-civil-airmen-stats.xlsx", sheet = 4, startRow = 6, na.strings = "N/Ap", cols = 1:11)
head(faa_raw)

str(faa_raw)
```

The hardest part is to identify the cells of the first column with bold 
formatting, indicating categories. I do this using the openxlsx package, then 
extract the category names from the relevant cells. That done, I use those
cells to fill the rest of the column. For simplicity's sake, I don't 
bother with subcategories (e.g., commercial fixed-wing vs. helicopter licenses)

After that, I clean up column names and delete an unnecessary row.
```{r}
faa_cleaned <- faa_raw[1:53, ] %>%
  mutate(category = NA_character_)
faa_cleaned[cells$row - 6, "category"] <- str_remove(cells$character, "\\s\\d.*|\\s?-.*") %>%
  str_trim()

faa_cleaned <- faa_cleaned %>%
  fill(category, .direction = "down") %>%
  rename_with(~ str_to_lower(.x) %>%
    str_replace_all("-?\\.", "_") %>%
    str_remove_all("_[^a-zA-Z]+")) %>%
  filter(!str_detect(class_of_certificate, "Total|^Airplane")) %>%
  mutate(str_remove(category, "\\sCetificates"))
```

From then on it's an easy pivoting job. 
```{r}
faa_cleaned <- faa_cleaned %>%
  select(-total) %>%
  pivot_longer(cols = alaskan:outside_u_s, names_to = "Region", values_to = "Licenses")
```

Now to actually plot the thing.
```{r}
p1 <- faa_cleaned %>%
  ggplot(aes(x = category, y = Licenses, fill = category)) +
  geom_col(position = "dodge") +
  facet_wrap(~Region) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_blank(),
    legend.text = element_text(size = 8),
    legend.title = element_blank(),
    legend.background = element_rect(fill = "khaki4", color = "gray93"),
    plot.background = element_rect(fill = "cadetblue3", color = "gold"),
    strip.text = element_text(color = "white"),
    strip.background = element_rect(fill = "black"),
    panel.background = element_rect(fill = "white")
  ) +
  scale_y_continuous(labels = scales::comma_format()) +
  scale_fill_brewer(palette = "Set3") +
  labs(x = "Category", title = "2019 Pilots' Licenses by Region and Category") +
  guides(fill = guide_legend(override.aes = list(color = "black"))) +
  theme(plot.title.position = "plot")
p1
ggsave("faa1.png", plot = p1, path = "../outputs/plots")
```
