---
title: "`r params$title`"
author: "Ryan Heslin"
date: "`r format(Sys.Date(), '%B %e, %Y')`"
header-includes:
  - \setlength{\parindent}{2em}
  - \setlength{\parskip}{2em}
params:
    title: "Hyperwar Analysis"
output:
    github_document:
      df_print: "kable"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  fig.pos = "",
  fig.path = file.path(
    "../figure",
    tools::file_path_sans_ext(
      knitr::current_input()
    ),
    "/"
  ),
  message = FALSE,
  tidy = "styler",
  warning = FALSE,
  fig.align = "center",
  highlight = TRUE
)
```

Today I'll tidy some data from [Hyperwar](https://www.ibiblio.org/hyperwar/AAF/StatDigest/aafsd-3.html), an
archive of World War II primary source documents. The data are several tables of
aircraft in Army Air Forces service each month of the war, broken down by role and type.
Challenges include nested headers, blank rows, and several nasty OCR errors.
```{r}
library(tidyverse)
library(rlang)
library(rvest)
library(lubridate)

url <- "https://www.ibiblio.org/hyperwar/AAF/StatDigest/aafsd-3.html"

sel <- "blockquote:nth-child(113) table , blockquote:nth-child(109) table, blockquote:nth-child(105) table, blockquote:nth-child(101) table"
```

Scraping is not too challenging.
```{r}

raw <- read_html(url) %>%
  html_elements(css = sel) %>%
  html_table()
```

After some experimentation, I devise a function to clean the data.
```{r}

clean_hyperwar <- function(tab) {
  tab <- tab[, colSums(is.na(tab)) != nrow(tab)]
  names(tab) <- c("Type", paste(tab[1, -1], names(tab)[-1]))
  tab %>%
    mutate(
      Category = str_extract(Type, "(^[^0-9\\-]+)(?=s--)"),
      .before = Category
    ) %>%
    fill(Category, .direction = "down") %>%
    mutate(
      Type = case_when(
        str_detect(Type, "2nd") ~ paste("2nd. Line", Category),
        str_detect(Type, "^Other") ~ paste("Other", Category),
        TRUE ~ Type
      )
    ) %>%
    mutate(
      across(
        -c(Type, Category),
        ~ str_replace_all(
          .x,
          c(
            "-" = "0", "(\\d+,\\d{3})\\d" = "\\1",
            "," = ""
          )
        ) %>%
          as.numeric()
      )
    ) %>%
    filter(
      if_all(-c(Type, Category), ~ !is.na(.x)) &
        !str_detect(Type, "Total|1st|^Combat Airplanes$")
    ) %>%
    pivot_longer(
      cols = -c(Type, Category),
      names_to = "Month-Year",
      values_to = "Number"
    ) %>%
    mutate(across(
      c(Type, Category),
      as.factor
    ))
}
```


Data in hand, I fill in missing combinations and recode an OCR glitch.
```{r}
cleaned <- map(raw, clean_hyperwar) %>%
  bind_rows() %>%
  mutate(
    `Month-Year` = my(`Month-Year`),
    Number = recode(Number, `7221` = 721)
  ) %>%
  complete(nesting(Category, Type), `Month-Year`, fill = list(Number = 0))
```

It seems trainers dominated inventory throughout the war, which makes sense.
```{r, message = FALSE}
my_theme()

breaks <- paste(rep(c("Jan", "Jun"), times = 4), rep(1941:1945, 2))

cleaned %>%
  group_by(`Month-Year`, Category) %>%
  summarize(Total = sum(Number)) %>%
  ggplot(aes(x = `Month-Year`, y = Total, fill = Category)) +
  stat_smooth(
    geom = "area", method = "loess", position = position_stack(),
    outline.type = "full",
    color = "black", alpha = .5, bandwith = .2, span = .2
  ) +
  scale_x_date(breaks = my(breaks), labels = breaks, expand = c(0, 0))
```

After filtering out trainers, the rising share of fighters is evidence of the
intensifying campaign for air superiority.
```{r, message=FALSE}
cleaned %>%
  filter(Category != "Trainer") %>%
  group_by(`Month-Year`, Category) %>%
  summarize(Total = sum(Number)) %>%
  ggplot(aes(x = `Month-Year`, y = Total, fill = Category)) +
  stat_smooth(
    geom = "area", method = "loess",
    position = position_stack(),
    outline.type = "full", color = "black", alpha = .5,
    bandwith = .2, span = .2
  ) +
  scale_x_date(breaks = my(breaks), labels = breaks, expand = c(0, 0))
```

I plot the top three types on hand each month.
These plots reveal a shift from fighters to heavy bombers and transports midway
through the war. There were initially some odd spikes I fixed by correcting OCR 
errors.
```{r}
top_types <- cleaned %>%
  group_by(`Month-Year`) %>%
  filter(Category != "Trainer" & !str_detect(Type, "2nd|Other")) %>%
  slice_max(n = 3, order_by = Number) %>%
  mutate(Type_ordered = tidytext::reorder_within(Type, Number, within = `Month-Year`)) %>%
  ungroup() %>%
  arrange(`Month-Year`) %>%
  mutate(
    Year = year(`Month-Year`), Month = month(`Month-Year`, label = TRUE),
    Type = fct_recode(Type, "C-47" = "C-47, C-53")
  )

pal <- RColorBrewer::brewer.pal(n = nlevels(top_types$Category), name = "Spectral") %>%
  setNames(levels(top_types$Category))
plots <- top_types %>%
  group_by(Year) %>%
  group_map(
    ~ ggplot(
      data = .x,
      aes(
        x = Type_ordered,
        label = str_replace(Type, "(.{4})(.+)", "\\1-\n\\2"),
        y = Number / 1000,
        fill = Category
      )
    ) +
      geom_col(position = "dodge") +
      geom_text(
        size = 3,
        vjust = -.2,
        fontface = "bold"
      ) +
      scale_fill_manual(values = pal) +
      tidytext::scale_x_reordered(expand = c(0, 0)) +
      facet_wrap(. ~ Month, nrow = 2, scales = "free") +
      scale_y_continuous(expand = expansion(mult = c(0, .3))) +
      labs(
        title = paste("Top Types by Month:", .y),
        x = NULL,
        y = "Number(Thousands)"
      ) +
      theme(
        legend.position = "bottom",
        legend.box.spacing = unit(0.05, "cm"),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
      )
  )

walk(plots, print)
```

This last plot shows the most-produced type each month, along with its proportion
of total aircraft received. The steady rise in production is obvious, as is 
the shift from fighters to heavy bombers.
```{r}
cleaned %>%
  group_by(`Month-Year`) %>%
  filter(Category != "Trainer" & !str_detect(Type, "2nd|Other|Liaison")) %>%
  mutate(Proportion = Number / sum(Number)) %>%
  slice_max(n = 1, order_by = Number) %>%
  ungroup() %>%
  arrange(`Month-Year`) %>%
  mutate(
    Year = year(`Month-Year`), Month = month(`Month-Year`, label = TRUE),
    Type = fct_recode(Type, "C-47" = "C-47, C-53")
  ) %>%
  ggplot(aes(x = `Month-Year`, y = Number, color = Category, label = Type, size = Proportion)) +
  geom_point() +
  geom_text(size = 3, nudge_y = -200, color = "black", face = "bold") +
  scale_size_continuous(labels = scales::percent_format()) +
  scale_x_date(breaks = my(breaks), labels = breaks) +
  theme(
    legend.position = "bottom",
    legend.box.spacing = unit(0.05, "cm")
  ) +
  labs(title = "Most Common Type by Month")
```
